Characterizing Signal Loss in the 21\,cm Reionization Power Spectrum: \\A Revised Study of PAPER-64

### 2018-09-10 PAPER.tex
version of paper and diffs sent to collaboration as response to referee report

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen. Because this signal is very faint and difficult to isolate, it is crucial to develop analysis techniques that maximize sensitivity and suppress contaminants in data. It is also imperative to understand the trade-offs between different analysis methods and their effects on power spectrum estimates. Specifically, with a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to understand how certain analysis choices can lead to the loss of the EoR signal. In this paper, we focus on signal loss associated with power spectrum estimation. We describe the origin of this loss using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of signal loss have led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. Additionally, we summarize errors associated with power spectrum error estimation that were previously unaccounted for. While we only focus on a subset of PAPER-64 data in this paper, revised power spectrum limits from the PAPER experiment are presented in a forthcoming paper by Kolopanis et al. (\textit{in prep.}) and supersede results from previously published PAPER analyses.
\end{abstract}

@@

Finally, we also highlight several additional errors made in previous PAPER analyses, including those related to bootstrapping and error estimation. This paper adds to the growing foundations of lessons which have been documented, for example, in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

@@

As a consequence of our investigations, we have also used a subset of PAPER-64 data to make a new power spectrum analysis. This serves as an illustrative example of using a signal injection framework, correctly computing errors via bootstrapping, and accurately estimating thermal noise. Our revised PAPER-64 limits are presented in Kolopanis et al. (\textit{in prep.}), which supersede all previously published PAPER limits. Because of the many challenges associated with signal loss and its estimation as described in this paper, Kolopanis et al. (\textit{in prep.}) use a straightforward power spectrum estimation approach that is not lossy. However, the main reasons for a previously underestimated limit and 
ways in which our new analysis differs can still be summarized by the following:

>>>>

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen. Because this signal is very faint and difficult to isolate, it is crucial to develop analysis techniques that maximize sensitivity and suppress contaminants in data. It is also imperative to understand the trade-offs between different analysis methods and their effects on power spectrum estimates. Specifically, with a statistical power spectrum detection in HERA's foreseeable future, it has become increasingly important to understand how certain analysis choices can lead to the loss of the EoR signal. In this paper, we focus on signal loss associated with power spectrum estimation. We describe the origin of this loss using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of signal loss have led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. Additionally, we summarize errors associated with power spectrum error estimation that were previously unaccounted for. We focus on a subset of PAPER-64 data in this paper; revised power spectrum limits from the PAPER experiment are presented in a forthcoming paper by Kolopanis et al. (\textit{in prep.}) and supersede results from previously published PAPER analyses.
\end{abstract}

@@

Finally, we also highlight several additional errors made in previous PAPER analyses, including those related to bootstrapping and error estimation. This paper accompanies the erratum of \citet{ali_et_al2018} and adds to the growing foundations of lessons which have been documented, for example, in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

@@

As a consequence of our investigations, we have also used a subset of PAPER-64 data to make a new power spectrum analysis. This serves as an illustrative example of using a signal injection framework, correctly computing errors via bootstrapping, and accurately estimating thermal noise. Our revised PAPER-64 limits are presented in Kolopanis et al. (\textit{in prep.}), which supersede all previously published PAPER limits. Because of the many challenges associated with signal loss and its estimation as described in this paper, Kolopanis et al. (\textit{in prep.}) use a straightforward power spectrum estimation approach that is not lossy. However, the main reasons for a previously underestimated limit (\citealt{ali_et_al2018}) and 
ways in which our new analysis differs can still be summarized by the following:
